
---
title: "MATH 216 Homework 3"
author: "WRITE YOUR NAME HERE"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)
library(Quandl)
library(knitr)
```






## Question 1:


Who  | Variables | Threshold | Test % Correct
------------- | ------------- | ------------- | -------------
Alden  | `job2 + height + diet3 + income2` | 0.5  | 84.1
Amanda | `income_level + job_new + body_type_buckets` | 0.5 | 70.7
Brenda | `income+height` | 0.5 | 83.2
James | `income + job + orientation + body_type` | 0.5 | 71.4
Albert | `height`| 0.5 | 83.0


Recall from `Lec12.R` and quiz

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
profiles <- read_csv(file="profiles.csv") %>% 
  filter(!is.na(height)) %>% 
  mutate(is_female=ifelse(sex=='f', 1, 0)) %>% 
  tibble::rownames_to_column(var="id")


profiles_clean <- profiles %>%
  filter(!is.na(height)) %>%
  filter(between(height, 55, 80))

prop_female_height <- profiles_clean %>%
  select(is_female, height) %>%
  group_by(height) %>%
  summarise(prop_female = mean(is_female))

# Plot
ggplot(data=prop_female_height, aes(height, prop_female)) +
  geom_line() + 
  geom_hline(yintercept = 0.5, col="red") +
  labs(y="Proportion Female")
```



## Question 2:

**Important Principle of Coding**: [Don't Repeat Yourself](https://en.wikipedia.org/wiki/Don%27t_repeat_yourself)

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# 1. Get gold & bitcoin data
# 2. Make column structure the same
# 3. Add variable type
gold <- Quandl("BUNDESBANK/BBK01_WT5511") %>% 
  select(Date, Value) %>% 
  mutate(type="Gold")

bitcoin <- Quandl("BAVERAGE/USD") %>% 
  rename(Value = `24h Average`) %>% 
  select(Date, Value) %>% 
  mutate(type="Bitcoin")

# Combine them into single data frame using bind_rows()
combined <- bind_rows(gold, bitcoin) %>% 
  # Group by here!
  group_by(type) %>% 
  # Then do the following ONLY ONCE:
  filter(year(Date) >= 2011) %>% 
  arrange(Date) %>% 
  mutate(
    Value_yest = lag(Value),
    rel_diff = 100 * (Value-Value_yest)/Value_yest
    )

# Plot
ggplot(combined, aes(x=Date, y=rel_diff, col=type)) +
  geom_line() +
  labs(y="% Change")
```





## Question 3:

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
jukebox <- read_csv(file="reed_jukebox.csv")

# Clean certain artists' names:
sigur_ros <- "Sigur Ro\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8c\x93\xa0\xbcs"
bjork <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk"
blue_oyster_cult <- "Blue O\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcyster Cult"
husker_do <- "Hu\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcsker Du\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbc"
bjork_brodsky <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk & Brodsky Quartet"
slagsmalsklubben <- "Slagsma\xfc\xbe\x99\x86\x94\xbc_lsklubben "
bjork_sugarcubes <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk (Sugarcubes)"

jukebox <- jukebox %>%
  mutate(
    artist = ifelse(artist == sigur_ros, "Sigor Ros", artist),
    artist = ifelse(artist == bjork, "Bjork", artist),
    artist = ifelse(artist == blue_oyster_cult, "Blue Oyster Cult", artist),
    artist = ifelse(artist == husker_do, "Husker Do", artist),
    artist = ifelse(artist == bjork_brodsky, "Bjork & Brodsky Quartet", artist),
    artist = ifelse(artist == slagsmalsklubben, "Slagsmalsklubben", artist),
    artist = ifelse(artist == bjork_sugarcubes, "Bjork (Sugarcubes)", artist)
  )
```


When parsing the time, what most of you did:

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
jukebox_hourly <- jukebox %>%
  mutate(
    date_time = parse_date_time(date_time, "%b %d %H%M%S %Y"),
    hour=hour(date_time)
    ) %>%
  group_by(hour) %>%
  summarise(count=n())
```


What's wrong with this plot?

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
ggplot(data=jukebox_hourly, aes(x=hour, y=count)) + 
  geom_bar(stat="identity") +
  xlab("Hour of day") + 
  ylab("Number of songs played")
```

Need to change the time zone using `date_time = with_tz(date_time, tz = "America/Los_Angeles")`:

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
jukebox_hourly <- jukebox %>%
  mutate(
    date_time = parse_date_time(date_time, "%b %d %H%M%S %Y"),
    date_time = with_tz(date_time, tz = "America/Los_Angeles"),
    hour=hour(date_time)
    ) %>%
  group_by(hour) %>%
  summarise(count=n())
ggplot(data=jukebox_hourly, aes(x=hour, y=count)) + 
  geom_bar(stat="identity") +
  xlab("Hour of day") + 
  ylab("Number of songs played")
```

Your top 10 graveyard shift artists are not

```{r, echo=FALSE}
top10_jukebox <- jukebox %>%
  # Cleaning the date variable
  mutate(
    clean_date = parse_date_time(date_time, "a b d HMS Y")
    ) %>%
  select(artist, clean_date) %>%
  # Filtering out months not in the academic year
  filter(month(clean_date) >= 9 | month(clean_date) <= 5) %>%
  # Filtering out times not in the graveyard shift
  filter(hour(clean_date) < 8 |
         hour(clean_date) == 8 & minute(clean_date) == 0 & second(clean_date) == 0) %>%
  # Group by artist and count
  group_by(artist) %>%
  count() %>%
  arrange(desc(n))

# Generating Top 10 Table:
head(top10_jukebox, 10) %>% kable()
```

but are instead below. Not that different actually!

```{r, echo=FALSE}
top10_jukebox <- jukebox %>%
  # Cleaning the date variable
  mutate(
    clean_date = parse_date_time(date_time, "a b d HMS Y"), 
    clean_date  = with_tz(clean_date , tz = "America/Los_Angeles")
    ) %>%
  select(artist, clean_date) %>%
  # Filtering out months not in the academic year
  filter(month(clean_date) >= 9 | month(clean_date) <= 5) %>%
  # Filtering out times not in the graveyard shift
  filter(hour(clean_date) < 8 |
         hour(clean_date) == 8 & minute(clean_date) == 0 & second(clean_date) == 0) %>%
  # Group by artist and count
  group_by(artist) %>%
  count() %>%
  arrange(desc(n))

# Generating Top 10 Table:
head(top10_jukebox, 10) %>% kable()
```





